---
title: "Quarto Homework Week 6"
author: "Ellie S Jones"
date: "October 5, 2026"

format:
    html:
      code-fold: true
      code-summary: "Click here to display code"
      theme: solar
---
```{r}
#| warning: false
#| message: false

## Load libraries
library(tidyverse)
library(here)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(fishualize)
library(praise)
```

## Water Chemistry Data
###### The following plot compares temperature and salinity by depth of water samples. 

```{r}
#| warning: false
#| message: false
#| out-width: "70%"
#| fig-align: "center"
#| label: fig-sal-temp-plot
#| fig-cap: "Water chemistry samples collected on January 15, 2021 by Nyssa Silbiger"

## Load data
CondData <- read.csv("CondData.csv")
DepthData <- read.csv("DepthData.csv")

## View data
view(CondData)
view(DepthData)

## Mutate and then join both datasets to compare temperature and salinity to depth
CondData <- CondData |>
  mutate(date = mdy_hms(date),
         date = round_date(date,"10 secs"))
DepthData <- DepthData |>
  mutate(date=ymd_hms(date))
JoinCondDepth <- CondData |>
  inner_join(DepthData,by="date")
view(JoinCondDepth)                    # Utilize joined conductivity and depth data from Week 5 homework

## Create a plot comparing salinity and temperature by depth
JCD_clean <- JoinCondDepth |>
  filter(Salinity > 30) |>
  filter(Depth > 0.2)                  # Filter out the outlier data points in salinity below 30 PSU and in depth above 0.2 so all data points are visible on the wrapped plots
JCD_clean |>
  pivot_longer(cols=c(Temperature,Salinity),
               names_to = "variable",
               values_to = "value") |>             # Pivot dataset longer to wrap in two plots
  ggplot(aes(x=value,y=Depth,color=variable)) +    # Create a data plot with depth against temperature and salinity
  geom_point() +
  facet_wrap(~variable, scales = "free_x",         # Wrap to create two plots
             labeller = as_labeller(c(
               Temperature = "Temperature (°C)",
               Salinity = "Salinity (PSU)"
             ))
             ) +
  scale_y_reverse() +                              # Reverse the y axis scale so the deepest depth is at the bottom and shallowest is at the top of the y axis
  labs(title = "Salinity and Temperature of Water Chemistry Samples by Depth",
       x = NULL,
       y = "Depth (m)")+
   scale_color_fish_d(option="Etheostoma_barrenense")+
  theme_bw()+
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position="none"
  )
```
Based on these data shown in @fig-sal-temp-plot, samples taken at slightly lower depths seem to be warmer and have slightly lower salinity. This could be a signal of submarine groundwater discharge or a freshwater seep, but I would expect that (based on my limited knowledge of SGD) to be colder and less saline. Due to the only slight variation in depth and variables, it is likely just showing slight shifts in the water as the samples were being collected.

###### The following table shows average temperature and salinity by depth of water samples.

```{r}
#| warning: false
#| message: false
#| label: tbl-sal-temp-table
#| tbl-cap: "Mean temperature and salinity by depth"

## Create a table
JCD_clean |>
  group_by(Depth) |>
  summarise("Mean Salinity (PSU)" = mean(Salinity, na.rm = TRUE), "Mean Temperature (°C)" = mean(Temperature, na.rm = TRUE)) |>
  kbl() |>                         # Use to turn data into tables using kableExtra
  kable_material_dark() |>         # Add a theme to the table
  column_spec(1,bold=TRUE) |>      # Bold the column headers
  kable_styling(bootstrap_options = 
                  c("hover","condensed"),
                full_width = FALSE)
```

#### Mahalo for reading my Quarto doc! 
