---
title: "Quarto Homework Week 6"
author: "Ellie S Jones"
date: "October 5, 2026"

format:
    html:
      code-fold: true
      code-summary: "Click here to display code"
      theme: minty
---
```{r}
#| warning: false
#| message: false

## Load libraries
library(tidyverse)
library(here)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(fishualize)
library(praise)
```

## Water Chemistry Data
###### The following plot compares temperature and salinity by depth of water samples. 

```{r}
#| warning: false
#| message: false
#| fig-align: "center"
#| label: fig-sal-temp-plot
#| fig-cap: "Water chemistry samples collected on January 15, 2021 by Nyssa Silbiger"

## Load data
CondData <- read_csv(here("Week_05","Data","CondData.csv"))
DepthData <- read.csv(here("Week_05","Data","DepthData.csv"))

## View data
view(CondData)

## Join data
CondData <- CondData |>
  mutate(date = mdy_hms(date),
         date = round_date(date,"10 secs"))
DepthData <- DepthData |>
  mutate(date=ymd_hms(date))
JoinCondDepth <- CondData |>
  inner_join(DepthData,by="date")
view(JoinCondDepth)        # Utilize joined conductivity and depth data from Week 5 homework

## Create a plot
JCD_clean <- JoinCondDepth |>
  filter(Salinity > 30) |>
  filter(Depth > 0.2) # Filter out the two outlier data points below 30 PSU
JCD_clean |>
  pivot_longer(cols=c(Temperature,Salinity),
               names_to = "variable",
               values_to = "value") |>      # Pivot dataset longer
  ggplot(aes(x=value,y=Depth,color=variable)) +    # Create a data plot with depth against temperature and salinity
  geom_point() +
  facet_wrap(~variable, scales = "free_x") +   # Wrap to create two plots
  scale_y_reverse() +   # Reverse the scale so the deepest depth is at the bottom and shallowest is at the top of the y axis
  labs(title = "Salinity and Temperature of Water Chemistry Samples by Depth",
       x = NULL,
       y = "Depth (m)")+
   scale_color_fish_d(option="Etheostoma_barrenense")+
  theme_minimal()+
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position="none"
  )
view(JCD_clean)

```
Based on these data, samples taken at slightly lower depths seem to be warmer and have slightly lower salinity (see @fig-sal-temp-plot). This could be indicative of a freshwater seep or other water input source in the location where these data were collected. 

###### The following table shows average temperature and salinity by depth of water samples.

```{r}
#| warning: false
#| message: false
#| label: tbl-sal-temp-table
#| tbl-cap: "Mean temperature and salinity by depth"

## Create a table
JCD_clean |>
  group_by(Depth) |>
  summarise(salmean = mean(Salinity, na.rm = TRUE), tempmean = mean(Temperature, na.rm = TRUE)) |>
  kbl() |>  # Use to turn data into tables using kableExtra
  kable_classic() |>  # Add a theme to the table
  column_spec(1,bold=TRUE) # Call out a specific row in the table

## I am very tired and it is very late but I enjoy coding very much
praise()
```

#### Mahalo for reading my Quarto doc! 
